### 2015-05-12 whoop.tex
Progress

A \emph{concurrent program} is described by a finite set of
\emph{shared} variables, $\sharedvars$, a finite set of mutexes,
$\mutexes$, and a finite set of \emph{thread templates}.  A thread
template, $T$, consists of a finite set of procedures,
$\procedures{T}$, and a finite set of private variables,
$\variables{T}$.  A designated procedure, procedure $\mainforthread{T}
\in \procedures{T}$, denotes the starting point for execution of a
template by a thread.  A procedure $P \in \procedures{T}$ is
represented by a control flow graph of basic blocks, where each block
contains a sequence of statements.  A basic block either has a single
successor or a pair of successors.  In the latter case, an \emph{exit
  condition} over thread-private variables determines, at runtime, the
successor to which control should flow on block exit.

@@

We consider a standard interleaving model of concurrency: at any
execution point, a thread may execute its current statement, unless
that statement has the form $\mutexlock(m)$ and mutex $m$ is already held
by some thread.  Executing a statement causes the state of the thread,
and the shared state, to be updated in a standard manner.  For
example, if a thread following template $T$ executes $s = e$, where $s
\in \sharedvars$ and $e$ is an expression over $\variables{T}$, the
shared variable valuation is updated so that $s$ has the value
determined by evaluating $e$ in the context of the thread's private
variable valuation.  The thread's next statement is updated according
to the control flow graph of the current procedure.

@@

\medskip\noindent\textbf{Lockset instrumentation} \ADComment{This is as far as I have gotten with the writing.  I now want to explain formally and somewhat abstractly how our lockset instrumentation works.}

>>>>

A \emph{concurrent program} is described by a finite set of
\emph{shared} variables, $\sharedvars$, a finite set of mutexes,
$\mutexes$, and a finite set of \emph{thread templates}.  A thread
template, $T$, consists of a finite set of procedures,
$\procedures{T}$, and a finite set of private variables,
$\variables{T}$.  A designated procedure, procedure $\mainforthread{T}
\in \procedures{T}$, denotes the starting point for execution of a
template by a thread.  A procedure in $\procedures{T}$ is
represented by a control flow graph of basic blocks, where each block
contains a sequence of statements.  A basic block either has a single
successor or a pair of successors.  In the latter case, an \emph{exit
  condition} over thread-private variables determines, at runtime, the
successor to which control should flow on block exit.

@@

We consider a standard interleaving model of concurrency: at any
execution point, a thread may execute its current statement, unless
that statement has the form $\mutexlock{m}$ and mutex $m$ is already held
by some thread.  Executing a statement causes the state of the thread,
and the shared state, to be updated in a standard manner.  For
example, if a thread following template $T$ executes $s = e$, where $s
\in \sharedvars$ and $e$ is an expression over $\variables{T}$, the
shared variable valuation is updated so that $s$ has the value
determined by evaluating $e$ in the context of the thread's private
variable valuation.  The thread's next statement is updated according
to the control flow graph of the current procedure.

@@

\medskip\noindent\textbf{Lockset instrumentation }
Suppose that $S$ and $T$ are thread templates for a concurrent program, including the possibly that $S=T$.
We wish to check whether it is possible for a thread that is an instance of $S$ to race with a thread that is
an instance of $T$.

We achieve this in the following manner.  Given an arbitrary symbol, $i$, we define the \emph{instrumentation} of a
template $T$ with respect to $i$, denoted $\instrument{T}{i}$.  There are two aspects to instrumentation, \emph{renaming} and \emph{lockset instrumentation}.

Renaming is simple: every private variable $x \in \variables{T}$ used in $T$ is replaced with a renamed variable $\instrument{x}{i}$ in $\instrument{T}{i}$, and every procedure $f \in \procedures{T}$ is renamed (both at declaration site and at all call sites) to $\instrument{f}{i}$ in $\instrument{T}{i}$.  The purpose of renaming is to ensure that when we analyse a pair of templates, $S$ and $T$, we ensure that $S$ and $T$ execute distinct procedures and operate on distinct private data.  Shared variables and mutexes are \emph{not} renamed, because these are indeed shared across templates.

\ADComment{Next up: describe lockset instrumentation with respect to Figure~\ref{ }.  This involves explaining what the sets $W$, $R$, $\mathit{LS}$ and $\mathit{CLS}$ are, and explaining how the instrumentation works.}

\ADComment{Finally, explain the main program that we check, which initializes the sets and runs the mains of each template.}

\ADComment{Last of all, pay attention to termination.}
